---
title: "Morris_meta"
author: "Weihsueh Chiu"
date: "February 9, 2017"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE, dpi=300}
library(Hmisc)
library(knitr)
library(pander)
library(metafor)

##### functions_for_effect_sizes

get_contin_es <- function(dat.contin, measure = "ROM", pct.scaling=TRUE, md.scaling=FALSE,
                          include.controls=FALSE) {
  endpoint.id.vec <- unique(dat.contin$endpoint.id)
  dat.contin.trans<-data.frame()
  for (endpoint.id.now in endpoint.id.vec) {
    dat.tmp <- subset(dat.contin,endpoint.id == endpoint.id.now) # & !Exclude)
    if (dim(dat.tmp)[1]>1) {
      dat.tmp$yi<-NA
      dat.tmp$vi<-NA
      control.indx<-dat.tmp$dose.index==0
      dose.index.vec<-dat.tmp$dose.index
      for (dose.indx in dose.index.vec) {
        treated.indx<-dat.tmp$dose.index==dose.indx
        if (pct.scaling) { # scaling to percent (0.01 -> 1)
          pct.scale <- c(1e2,1e4)
        } else {
          pct.scale <- c(1,1)
        }
        # Log ratio of mean difference, multiplied by 100 (variance multiplied by 10,000)
        if (measure == "MD" ) { # Mean difference, with optional scaling by control value
          if (md.scaling) { 
            md.scale <- c(dat.tmp$response[control.indx],dat.tmp$response[control.indx]^2)
          } else {
            md.scale <- c(1,1)
          }
          dat.tmp[treated.indx,c("yi","vi")]<-
            escalc(measure="MD",m2i=dat.tmp$response[control.indx],m1i=dat.tmp$response[treated.indx],
                   sd2i=dat.tmp$stdev[control.indx],sd1i=dat.tmp$stdev[treated.indx],
                   n2i=dat.tmp$N[control.indx],n1i=dat.tmp$N[treated.indx])/md.scale*pct.scale
        }
        else { # default is ROM
          dat.tmp[treated.indx,c("yi","vi")]<-
            escalc(measure="ROM",m2i=dat.tmp$response[control.indx],m1i=dat.tmp$response[treated.indx],
                   sd2i=dat.tmp$stdev[control.indx],sd1i=dat.tmp$stdev[treated.indx],
                   n2i=dat.tmp$N[control.indx],n1i=dat.tmp$N[treated.indx])*pct.scale
        } 
      }
      # Do not include controls
      if (include.controls) {
        dat.contin.trans<-rbind(dat.contin.trans,dat.tmp)
      } else {
        dat.contin.trans<-rbind(dat.contin.trans,dat.tmp[!control.indx,])
      }
    }
  }    
  # return data frame
  dat.contin.trans
}

```

```{r more_functions, echo=FALSE}
##### forestfunction

doforest <- function(dat.forest,effectlab="",main="",...) {
  dosepos<-max(dat.forest$yi+3*sqrt(dat.forest$vi),na.rm=TRUE)
  dat.forest.rma<- rma(yi,vi=vi,slab=slab,data=dat.forest)
  print(summary(dat.forest.rma))
  forest(dat.forest.rma,xlab=effectlab
         ,order=order(dat.forest$brom,
                      dat.forest$dose,dat.forest$study.name),ilab=cbind(dat.forest$dose),
         ilab.xpos=dosepos,...)
  text(dat.forest.rma$ci.ub,-1,pos=4,paste("(I2=",round(dat.forest.rma$I2,1),"%)",sep=""),
       ...)
  mtext("Study and animal group",side=3,line=-1,adj=0,...)
  mtext("Estimate [95% CI]",side=3,line=-1,adj=1,...)
  mtext("Dose (mg/kg-d)",side=3,line=-1,at=dosepos,...)
  mtext(main,side=3,cex=1)
  dat.forest.rma
}


doforestdose <- function(dat.forest,effectlab="",main="",...) {
  dosepos<-max(dat.forest$yi+3*sqrt(dat.forest$vi),na.rm=TRUE)
  dat.forest.rma.dose<- rma(yi,vi=vi,slab=slab,data=dat.forest,mods = ~log10(dose))
  print(summary(dat.forest.rma.dose))
  forest(dat.forest.rma.dose,xlab=effectlab
         ,order=order(dat.forest$dose,dat.forest$study.name),ilab=cbind(dat.forest$dose),
         ilab.xpos=dosepos,...)
  text(dat.forest.rma.dose$ci.ub[2],-1,pos=4,paste(
    "(R2=",round(dat.forest.rma.dose$R2,1),"%, ",
    "I2=",round(dat.forest.rma.dose$I2,1),"%)",sep=""),
    ...)
  #        addpoly(dat.forest.rma.dose$b[1,1],sei=dat.forest.rma.dose$se[1],
  #                mlab=paste("Intercept (",effectlab," at 1 mg/kg-d)",sep=""),cex=cex,row=-1
  #                ,...
  #                )
  addpoly(dat.forest.rma.dose$b[2,1],sei=dat.forest.rma.dose$se[2],
          mlab=paste("Slope (",effectlab,"\n per log10 change in exposure)",sep=""),row=-1
          ,...
  )
  mtext("Study and animal group",side=3,line=-1,adj=0,...)
  mtext("Estimate [95% CI]",side=3,line=-1,adj=1,...)
  mtext("Dose (mg/kg-d)",side=3,line=-1,at=dosepos,...)
  mtext(main,side=3,cex=1)
  dat.forest.rma.dose
}

##### forest dose linear no intercept default

doforestdose1 <- function(dat.forest,intercept=FALSE,scale=1,
                          effectlab="",main="",...) {
  dosepos<-max(dat.forest$yi+3*sqrt(dat.forest$vi),na.rm=TRUE)
  dat.forest$dose10 <- dat.forest$dose/10
  if (!intercept) {
    dat.forest.rma.dose<- rma(yi,vi=vi,slab=slab,data=dat.forest,mods = ~dose10-1)
  } else {
    dat.forest.rma.dose<- rma(yi,vi=vi,slab=slab,data=dat.forest,mods = ~dose10)
  }
  print(summary(dat.forest.rma.dose))
  forest(dat.forest.rma.dose,xlab=effectlab
         ,order=order(dat.forest$dose,dat.forest$study.name),ilab=cbind(dat.forest$dose),
         ilab.xpos=dosepos,...)
  text(dat.forest.rma.dose$ci.ub[1],-1,pos=4,paste(
    "(I2=",round(dat.forest.rma.dose$I2,1),"%)",sep=""),
    ...)
  if (intercept) {
    addpoly(dat.forest.rma.dose$b[1,1],sei=dat.forest.rma.dose$se[1],
            mlab=paste("Intercept",sep=""),row=-1
            ,...
    )
    roffset<-1
  } else {
    roffset<-0
  }
  addpoly(scale*dat.forest.rma.dose$b[roffset+1,1],sei=scale*dat.forest.rma.dose$se[roffset+1],
          mlab=paste("Linear (",effectlab," per\n ",scale," mg/kg-d)",sep=""),row=-1-roffset
          ,...
  )
  mtext("Study and animal group",side=3,line=-1,adj=0,...)
  mtext("Estimate [95% CI]",side=3,line=-1,adj=1,...)
  mtext("Dose (mg/kg-d)",side=3,line=-1,at=dosepos,...)
  mtext(main,side=3,cex=1)
  dat.forest.rma.dose
}

##### forest dose quadratic no intercept default

doforestdose2 <- function(dat.forest,intercept=FALSE,scale=1,
                          effectlab="",main="",...) {
  dosepos<-max(dat.forest$yi+3*sqrt(dat.forest$vi),na.rm=TRUE)
  dat.forest$dose10 <- dat.forest$dose/10
  if (!intercept) {
    dat.forest.rma.dose<- rma(yi,vi=vi,slab=slab,data=dat.forest,mods = ~dose10+I(dose10^2)-1)
  } else {
    dat.forest.rma.dose<- rma(yi,vi=vi,slab=slab,data=dat.forest,mods = ~dose10+I(dose10^2))
  }
  print(summary(dat.forest.rma.dose))
  forest(dat.forest.rma.dose,xlab=effectlab
         ,order=order(dat.forest$dose,dat.forest$study.name),ilab=cbind(dat.forest$dose),
         ilab.xpos=dosepos,...)
  if (is.null(dat.forest.rma.dose$R2)) {
    text(dat.forest.rma.dose$ci.ub[2],-1,pos=4,paste(
      "(I2=",round(dat.forest.rma.dose$I2,1),"%)",sep=""),
      ...)
  } else {
    text(dat.forest.rma.dose$ci.ub[2],-1,pos=4,paste(
      "(R2=",round(dat.forest.rma.dose$R2,1),"%, ",
      "I2=",round(dat.forest.rma.dose$I2,1),"%)",sep=""),
      ...)
  }
  if (intercept) {
    addpoly(dat.forest.rma.dose$b[1,1],sei=dat.forest.rma.dose$se[1],
            mlab=paste("Intercept",sep=""),row=-1
            ,...
    )
    roffset<-1
  } else {
    roffset<-0
  }
  addpoly(scale*dat.forest.rma.dose$b[roffset+1,1],sei=scale*dat.forest.rma.dose$se[roffset+1],
          mlab=paste("Linear (",effectlab," per ",scale," mg/kg-d)",sep=""),row=-1-roffset
          ,...
  )
  if (dat.forest.rma.dose$se[roffset+2] !=0) {
    addpoly(scale^2*dat.forest.rma.dose$b[roffset+2,1],sei=scale^2*dat.forest.rma.dose$se[roffset+2],
            mlab=paste("Quadratic (",effectlab,"\n per [",scale," mg/kg-d]^2)",sep=""),row=-2-roffset
            ,...
    )
  }
  mtext("Study and animal group",side=3,line=-1,adj=0,...)
  mtext("Estimate [95% CI]",side=3,line=-1,adj=1,...)
  mtext("Dose (mg/kg-d)",side=3,line=-1,at=dosepos,...)
  mtext(main,side=3,cex=1)
  dat.forest.rma.dose
}


##### doseresponsefunction

dodoseplot <- function(dat.forest,effectlab="",ylim=c(-100,100),logxlim=NA,main="") {
  if (is.na(logxlim[1])) {
    x1 <- floor(log10(range(dat.forest$dose)))
  } else {
    x1 <- logxlim
  }
  xlim<-10^(x1+c(0,1))
  
  errbar(dat.forest$dose,dat.forest$yi,dat.forest$yi+sqrt(dat.forest$vi)*qnorm(0.95),
         dat.forest$yi+sqrt(dat.forest$vi)*qnorm(0.05),
         xlab=paste("Dose mg/kg-d"),xlim=xlim,
         ylab=effectlab,axes=FALSE,log="x",ylim=ylim)
  abline(0,0,col="grey")
  box()
  axis(2)
  x2 <- seq(x1[1], x1[2]+1)
  ticksat <- as.vector(sapply(x2, function(p) (1:10)*10^p))
  axis(1, 10^x2)
  axis(1, ticksat, labels=NA, tcl=-0.25, lwd=0, lwd.ticks=1)
  mtext(main,side=3)
  print("")
}


adddosepredict <- function(dat.now,dat.now.rma.dose,col.ci="red",col.cr="blue",lwd=2,lty.m=1,lty.ci=2,...) {
  xr <- log10(range(dat.now$dose))
  xr[1]<-min(xr,0)
  xx <- 10^seq(from=xr[1],to=xr[2],length.out=1000)
  yypred<-predict(dat.now.rma.dose,newmods=log10(xx))
  lines(xx,yypred$pred,lty=lty.m,col=col.ci,lwd=lwd,...)
  lines(xx,yypred$ci.lb,lty=lty.ci,col=col.ci,lwd=lwd,...)
  lines(xx,yypred$ci.ub,lty=lty.ci,col=col.ci,lwd=lwd,...)
  lines(xx,yypred$cr.lb,lty=lty.ci,col=col.cr,lwd=lwd+1,...)
  lines(xx,yypred$cr.ub,lty=lty.ci,col=col.cr,lwd=lwd+1,...)
}

adddose2predict <- function(dat.now,dat.now.rma.dose,col.ci="red",col.cr="blue",lwd=2,lty.m=1,lty.ci=2,
                            dosescale=1,...) {
  xr <- log10(range(dat.now$dose))
  xr[1]<-min(xr,0)
  xx <- (10^seq(from=xr[1],to=xr[2],length.out=1000))/dosescale
  yypred<-predict(dat.now.rma.dose,newmods=cbind(xx,xx^2))
  lines(dosescale*xx,yypred$pred,lty=lty.m,col=col.ci,lwd=lwd,...)
  lines(dosescale*xx,yypred$ci.lb,lty=lty.ci,col=col.ci,lwd=lwd,...)
  lines(dosescale*xx,yypred$ci.ub,lty=lty.ci,col=col.ci,lwd=lwd,...)
  lines(dosescale*xx,yypred$cr.lb,lty=lty.ci,col=col.cr,lwd=lwd+1,...)
  lines(dosescale*xx,yypred$cr.ub,lty=lty.ci,col=col.cr,lwd=lwd+1,...)
}

## Linear model
adddose1predict <- function(dat.now,dat.now.rma.dose,col.ci="red",col.cr="blue",lwd=2,lty.m=1,lty.ci=2,
                            dosescale=1,...) {
  xr <- log10(range(dat.now$dose))
  xr[1]<-min(xr,0)
  xx <- (10^seq(from=xr[1],to=xr[2],length.out=1000))/dosescale
  yypred<-predict(dat.now.rma.dose,newmods=xx)
  lines(dosescale*xx,yypred$pred,lty=lty.m,col=col.ci,lwd=lwd,...)
  lines(dosescale*xx,yypred$ci.lb,lty=lty.ci,col=col.ci,lwd=lwd,...)
  lines(dosescale*xx,yypred$ci.ub,lty=lty.ci,col=col.ci,lwd=lwd,...)
  lines(dosescale*xx,yypred$cr.lb,lty=lty.ci,col=col.cr,lwd=lwd+1,...)
  lines(dosescale*xx,yypred$cr.ub,lty=lty.ci,col=col.cr,lwd=lwd+1,...)
}

addbmd <- function(dat.now,dat.now.rma.dose,bmr=100*log(1.05),col="grey",line=-1,...) {
  xr <- log10(range(dat.now$dose))
  xr[1]<-min(xr,0)
  xx <- seq(from=xr[2],to=xr[1],length.out=1000) # go down from higher dose
  yypred<-predict(dat.now.rma.dose,newmods=xx)
  if (bmr>0) {
    bmd<-data.frame(m=10^approx(yypred$pred,xx,bmr)$y,
                    ci.lb=10^approx(yypred$ci.ub,xx,bmr)$y,
                    ci.ub=10^approx(yypred$ci.lb,xx,bmr)$y
    )
  } else {
    bmd<-data.frame(m=10^approx(yypred$pred,xx,bmr)$y,
                    ci.lb=10^approx(yypred$ci.lb,xx,bmr)$y,
                    ci.ub=10^approx(yypred$ci.ub,xx,bmr)$y
    )
  }
  #  points(bmd$m,bmr,pch=15,col=col)
  #  lines(c(bmd$ci.lb,bmd$ci.ub),rep(bmr,2),lwd=3,col=col,...)
  lines(c(1e-30,bmd$ci.lb,bmd$m,bmd$ci.ub),rep(bmr,4),lty=2,col="grey")
  lines(rep(bmd$ci.lb,2),c(bmr,-1e6),lty=2,col="grey")
  lines(rep(bmd$ci.ub,2),c(bmr,-1e6),lty=2,col="grey")
  lines(rep(bmd$m,2),c(bmr,-1e6),col="grey")
  mtext(paste("BMD(",signif(bmr,2),")=",signif(bmd$m,3),"[",signif(bmd$ci.lb,3),", ",
              signif(bmd$ci.ub,3),"]",sep=""),side=1,line=line,cex=0.8,
        at=bmd$m)
  bmd$bmr <- bmr
  bmd
}

addbmd2 <- function(dat.now,dat.now.rma.dose,bmr=100*log(1.05),col="grey",line=-1,dosescale=1,...) {
  xr <- log10(range(dat.now$dose))
  xr[1]<-min(xr,0)
  xx <- 10^seq(from=xr[1],to=xr[2],length.out=1000)/dosescale # go up from lower dose
  yypred<-predict(dat.now.rma.dose,newmods=cbind(xx,xx^2))
  if (bmr>0) {
    bmd<-data.frame(m=10^approx(yypred$pred,log10(xx),bmr)$y,
                    ci.lb=10^approx(yypred$ci.ub,log10(xx),bmr)$y,
                    ci.ub=10^approx(yypred$ci.lb,log10(xx),bmr)$y
    )
  } else {
    bmd<-data.frame(m=10^approx(yypred$pred,log10(xx),bmr)$y,
                    ci.lb=10^approx(yypred$ci.lb,log10(xx),bmr)$y,
                    ci.ub=10^approx(yypred$ci.ub,log10(xx),bmr)$y
    )
  }
  bmd<-bmd*dosescale
  #  points(bmd$m,bmr,pch=15,col=col)
  #  lines(c(bmd$ci.lb,bmd$ci.ub),rep(bmr,2),lwd=3,col=col,...)
  lines(c(1e-30,bmd$ci.lb,bmd$m,bmd$ci.ub),rep(bmr,4),lty=2,col="grey")
  lines(rep(bmd$ci.lb,2),c(bmr,-1e6),lty=2,col="grey")
  lines(rep(bmd$ci.ub,2),c(bmr,-1e6),lty=2,col="grey")
  lines(rep(bmd$m,2),c(bmr,-1e6),col="grey")
  mtext(paste("BMD(",signif(bmr,2),")=",signif(bmd$m,3),"[",signif(bmd$ci.lb,3),", ",
              signif(bmd$ci.ub,3),"]",sep=""),side=1,line=line,cex=0.8,
        at=bmd$m)
  bmd$bmr <- bmr
  bmd
}

addbmd1 <- function(dat.now,dat.now.rma.dose,bmr=100*log(1.05),col="grey",line=-1,dosescale=1,...) {
  xr <- log10(range(dat.now$dose))
  xr[1]<-min(xr,0)
  xx <- 10^seq(from=xr[1],to=xr[2],length.out=1000)/dosescale # go up from lower dose
  yypred<-predict(dat.now.rma.dose,newmods=xx)
  if (bmr>0) {
    bmd<-data.frame(m=10^approx(yypred$pred,log10(xx),bmr)$y,
                    ci.lb=10^approx(yypred$ci.ub,log10(xx),bmr)$y,
                    ci.ub=10^approx(yypred$ci.lb,log10(xx),bmr)$y
    )
  } else {
    bmd<-data.frame(m=10^approx(yypred$pred,log10(xx),bmr)$y,
                    ci.lb=10^approx(yypred$ci.lb,log10(xx),bmr)$y,
                    ci.ub=10^approx(yypred$ci.ub,log10(xx),bmr)$y
    )
  }
  bmd<-bmd*dosescale
  #  points(bmd$m,bmr,pch=15,col=col)
  #  lines(c(bmd$ci.lb,bmd$ci.ub),rep(bmr,2),lwd=3,col=col,...)
  lines(c(1e-30,bmd$ci.lb,bmd$m,bmd$ci.ub),rep(bmr,4),lty=2,col="grey")
  lines(rep(bmd$ci.lb,2),c(bmr,-1e6),lty=2,col="grey")
  lines(rep(bmd$ci.ub,2),c(bmr,-1e6),lty=2,col="grey")
  lines(rep(bmd$m,2),c(bmr,-1e6),col="grey")
  mtext(paste("BMD(",signif(bmr,2),")=",signif(bmd$m,3),"[",signif(bmd$ci.lb,3),", ",
              signif(bmd$ci.ub,3),"]",sep=""),side=1,line=line,cex=0.8,
        at=bmd$m)
  bmd$bmr <- bmr
  bmd
}



getresults.sum <- function(dat.rma,b.lab="",bindx=1) {
	z<- data.frame(effect=effectlab,b.lab=b.lab,
						b.parm=row.names(dat.rma$b)[bindx],
						b=dat.rma$b[bindx], 
						ci.lb=dat.rma$ci.lb[bindx], ci.ub=dat.rma$ci.ub[bindx], pval=dat.rma$pval[bindx],
						tau=sqrt(dat.rma$tau2),I2=dat.rma$I2,QEp=dat.rma$QEp,
						AICc=dat.rma$fit.stats["AICc","REML"])
	z
}

getbmd.results.sum <- function(dat.bmd,b.lab="") {
	z<- data.frame(effect=effectlab,b.lab=b.lab,
					bmr=dat.bmd$bmr,bmd=dat.bmd$m,ci.lb=dat.bmd$ci.lb,ci.ub=dat.bmd$ci.ub)
	z
}


```

## Morris Latency last trial animal details

Only included studies that had standard deviation data extracted.  Effect measure is log ratio of mean between treated and control, times 100 (which for small values is close to the % change).  Random effects models were fit for overall effect, linear in log10(dose), linear in dose, and linear-quadratic in dose.  Sensitivity analysis included leaving one study out at a time and using only the highest dose from each study.

The results show a statistically significant overall effect of PBDE treatment that is also robust to leaving out individual studies, using only the highest dose group in each study, and leaving out individual studies using the highest dose group only.  There was also low or no heterogeneity (<25%, not statistically significant in primary analysis and all sensitivity analyses).  Use of a linear model in log10(dose), a linear model in dose, or a linear-quadratic model did not reduce heterogeneity, and in some cases increased it.  Thus, there is no evidence of a dose-response gradient.

```{r calc_effect_sizes, fig.width=6.5, fig.height = 8, echo=FALSE, dpi=300}
##### calc_effect_sizes

Morris.dat <- read.delim("PBDE_Morris_latency_last_or_avail_trial_inclall.dat",sep="\t",header=TRUE,as.is=TRUE)
Morris.dat.incl <- subset(Morris.dat,!Exclude & !is.na(response))
#Morris.dat.incl$stdev[is.na(Morris.dat.incl$stdev)] <- 1000
Morris.dat.trans<- get_contin_es(Morris.dat.incl)
# plot(Morris.dat.trans$yi,Morris.dat.trans$percent.control.mean,pch=15,xlab="log Ratio of mean (x100)",ylab="HAWC % difference")

Morris.dat.trans$brom <- as.numeric(t(as.data.frame(strsplit(Morris.dat.trans$chemical," "))[2,]))
Morris.dat.trans$brom[as.vector((as.data.frame(strsplit(Morris.dat.trans$chemical,
                                                        " "))[1,])=="DE")]<-0
dat.order<-Morris.dat.trans[order(Morris.dat.trans$brom,Morris.dat.trans$dose),]
dat.order$vi[is.na(dat.order$vi)]<-0
dat.order$col<-"black"
dat.order$col[dat.order$vi==0]<-"royalblue3"
dat.order$slab <- paste(dat.order$study.name,"\n  ",dat.order$chemical,dat.order$animal.group.name)
dosepos<-max(dat.order$yi+3*sqrt(dat.order$vi),na.rm=TRUE)

# pdf("Fig5.pdf",height=8,width=6.5)
forest(x=dat.order$yi,vi=dat.order$vi,slab=dat.order$slab,col=dat.order$col,
       xlab="Latency last trial log(Ratio of mean)x100",main="PBDE All Studies and Doses",
       ilab=cbind(dat.order$dose), ilab.xpos=dosepos,
       cex=0.5,psize=2)
abline(h=(1:(length(dat.order$brom)-1))[rev(diff(dat.order$brom))!=0]+0.55,lty=2)
mtext("Study, congener, and animal group",side=3,line=-1,adj=0,cex=0.7)
mtext("Estimate [95% CI]",side=3,line=-1,adj=1,cex=0.7)
mtext("Dose (mg/kg-d)",side=3,line=-1,at=dosepos,cex=0.7)
# dev.off()
```


```{r Morris Latency last trial animal, fig.width=6.5, fig.height = 8, echo=FALSE, dpi=300}
##### Morris Latency_animal_forest_dose

results.sum <- data.frame()
bmd.results.sum<-data.frame()

dat<-Morris.dat.trans
effectlab<-"Latency last trial log(Ratio of mean)x100"
dat$slab <- paste(dat$study.name,"\n  ",dat$chemical,dat$animal.group.name)
      # All doses for each study
      dat.now <- dat
      studylist <- unique(dat.now$study.name)
      nstudy<-length(studylist)
            # Intercept only
        print("Overall Effect---------------------")
        
# pdf("Fig6.pdf",height=8,width=6.5)
        dat.now.rma <- doforest(dat.now,effectlab=effectlab,
                                main=paste("PBDE All Doses"),cex=0.7)
        dat.order<-dat.now[order(dat.now$brom,
                      dat.now$dose,dat.now$study.name),]
        dat.order.nona<-dat.order[!is.na(dat.order$vi),]
        abline(h=(1:(length(dat.order.nona$brom)-1))[rev(diff(dat.order.nona$brom))!=0]+0.5,lty=2)
		results.sum <- rbind(results.sum,getresults.sum(dat.now.rma,b.lab="Overall",bindx=1))
        # dev.off()
        #        print(dat.now.rma.l1o<-leave1out(dat.now.rma))
        # Leave one out
        print("Leave one out---------------------")
        studylist <- unique(dat.now$study.name)
        nstudy<-length(studylist)
        dat.loo.rma.list<-list()
        if (nstudy>2) { 
          for (n in 1:nstudy) {
            study.now<-studylist[n]
            dat.loo <- subset(dat.now,study.name != study.now)
            print(dat.loo.rma.list[[n]]<-rma(yi,vi=vi,slab=slab,data=dat.loo))
			results.sum <- rbind(results.sum,getresults.sum(dat.loo.rma.list[[n]],
				b.lab=paste("Overall minus",study.now),bindx=1))
            #print(predict(dat.loo.rma.list[[n]]))
            # forest(dat.loo$yi.back,ci.lb=dat.loo$ci.lb.back,ci.ub=dat.loo$ci.ub.back,slab=dat.loo$slab,
            #                             xlab=effectlab,xlim=c(xmin,xmax),at=xtx,
            #                             ilab=cbind(dat.loo$dose),
            #                             ilab.xpos=dosepos,cex=cex.lab,
            #                             rows=rnums,ylim=c(-1,ymax),
            #                             digits=2
            # )
          }
        }
        # Intercept and log10(dose) - test for trend
        print("Linear in log10(dose)---------------------")
        dat.now.rma.dose <- doforestdose(dat.now,ylim=c(-1.5,dim(dat.now)[1]+3),
                                         effectlab=effectlab,
                                         main=paste("PBDE All Doses"),cex=0.7)
		results.sum <- rbind(results.sum,getresults.sum(dat.now.rma.dose,b.lab="Trend in log10(dose)",bindx=2))
        # Linear - no intercept - for estimating BMD
        print("Linear in dose---------------------")
        dat.now.rma.dose1 <- doforestdose1(dat.now,ylim=c(-2.5,dim(dat.now)[1]+3),
                                           effectlab=effectlab,
                                           main=paste("PBDE All Doses"),cex=0.7)
		results.sum <- rbind(results.sum,getresults.sum(dat.now.rma.dose1,b.lab="Linear in dose10",bindx=1))
        # Quadratic - no intercept - for estimating BMD
        print("Linear or LinearQuadratic in dose---------------------")
        dat.now.rma.dose2 <- doforestdose2(dat.now,ylim=c(-2.5,dim(dat.now)[1]+3),
                                           effectlab=effectlab,
                                           main=paste("PBDE All Doses"),cex=0.7)
		results.sum <- rbind(results.sum,getresults.sum(dat.now.rma.dose2,b.lab="LinearQuadratic in dose10",bindx=1))
		results.sum <- rbind(results.sum,getresults.sum(dat.now.rma.dose2,b.lab="LinearQuadratic in dose10",bindx=2))
        par(mfrow=c(3,1))
        dodoseplot(dat.now,effectlab=effectlab,logxlim=c(-2,1))
        adddosepredict(dat.now,dat.now.rma.dose)
        mtext("Log-linear model",side=3,line=-1,cex=0.8)
        dodoseplot(dat.now,effectlab=effectlab,logxlim=c(-2,1))
        adddose1predict(dat.now,dat.now.rma.dose1,dosescale=10)
        print(dat.now.bmd1<-addbmd1(dat.now,dat.now.rma.dose1,dosescale=10))
		bmd.results.sum <- rbind(bmd.results.sum,getbmd.results.sum(dat.now.bmd1,b.lab="Linear in dose10"))
        mtext("Linear model",side=3,line=-1,cex=0.8)
        dodoseplot(dat.now,effectlab=effectlab,logxlim=c(-2,1))
        adddose2predict(dat.now,dat.now.rma.dose2,dosescale=10)
        print(dat.now.bmd2<-addbmd2(dat.now,dat.now.rma.dose2,dosescale=10))
		bmd.results.sum <- rbind(bmd.results.sum,getbmd.results.sum(dat.now.bmd2,b.lab="LinearQuadratic in dose10"))
        mtext("Linear-quadratic model",side=3,line=-1,cex=0.8)
        par(mfrow=c(1,1))
        # dev.off()

        
      # highest dose for each study
      print(paste(effectlab,"Highest Dose"))
      dat.now <- subset(dat,high_dose==dose)
        # Intercept only
        # pdf(file=paste(phthalate.now,"HighestDose",species.now,"AGD.V5.pdf",sep="."),height=8,width=6.5)
        print("Highest Dose Overall Effect---------------------")
        dat.now.rma <- doforest(dat.now,effectlab=effectlab,
                                main=paste("PBDE Highest Doses"),cex=0.7)
		results.sum <- rbind(results.sum,getresults.sum(dat.now.rma,b.lab="Highest Doses-Overall",bindx=1))
		
		 # Leave one out
        print("Highest Dose Leave one out---------------------")
        studylist <- unique(dat.now$study.name)
        nstudy<-length(studylist)
        dat.loo.rma.list<-list()
        if (nstudy>2) { 
          for (n in 1:nstudy) {
            study.now<-studylist[n]
            dat.loo <- subset(dat.now,study.name != study.now)
            print(dat.loo.rma.list[[n]]<-rma(yi,vi=vi,slab=slab,data=dat.loo))
			results.sum <- rbind(results.sum,getresults.sum(dat.loo.rma.list[[n]],
				b.lab=paste("Overall minus",study.now),bindx=1))
            #print(predict(dat.loo.rma.list[[n]]))
            # forest(dat.loo$yi.back,ci.lb=dat.loo$ci.lb.back,ci.ub=dat.loo$ci.ub.back,slab=dat.loo$slab,
            #                             xlab=effectlab,xlim=c(xmin,xmax),at=xtx,
            #                             ilab=cbind(dat.loo$dose),
            #                             ilab.xpos=dosepos,cex=cex.lab,
            #                             rows=rnums,ylim=c(-1,ymax),
            #                             digits=2
            # )
          }
        }
        #        print(dat.now.rma.l1o<-leave1out(dat.now.rma))
        # Linear - no intercept - for estimating BMD
        print("Highest Dose Linear or LinearQuadratic in dose---------------------")
        dat.now.rma.dose1 <- doforestdose1(dat.now,ylim=c(-2.5,dim(dat.now)[1]+3),
                                           effectlab=effectlab,
                                           main=paste("PBDE Highest Doses"),cex=0.7)
		results.sum <- rbind(results.sum,getresults.sum(dat.now.rma.dose1,b.lab="Highest Doses-Linear in dose10",bindx=1))
        # Intercept and log10(dose) - test for trend
          print("Highest Dose Linear in log10(dose)---------------------")
          dat.now.rma.dose <- doforestdose(dat.now,ylim=c(-1.5,dim(dat.now)[1]+3),
                                           effectlab=effectlab,
                                           main=paste("PBDE Highest Doses"),cex=0.7)
		results.sum <- rbind(results.sum,getresults.sum(dat.now.rma.dose,b.lab="Highest Doses-Trend in log10(dose)",bindx=2))
          # Quadratic - no intercept - for estimating BMD
          print("Highest Dose LinearQuadratic in dose---------------------")
          dat.now.rma.dose2 <- doforestdose2(dat.now,ylim=c(-2.5,dim(dat.now)[1]+3),
                                             effectlab=effectlab,
                                             main=paste("PBDE Highest Doses"),cex=0.7)
		results.sum <- rbind(results.sum,getresults.sum(dat.now.rma.dose2,b.lab="Highest Doses-LinearQuadratic in dose10",bindx=1))
		results.sum <- rbind(results.sum,getresults.sum(dat.now.rma.dose2,b.lab="Highest Doses-LinearQuadratic in dose10",bindx=2))
          par(mfrow=c(3,1))
          dodoseplot(dat.now,effectlab=effectlab,logxlim=c(-2,1))
          adddosepredict(dat.now,dat.now.rma.dose)
          mtext("Log-linear model",side=3,line=-1,cex=0.8)
          dodoseplot(dat.now,effectlab=effectlab,logxlim=c(-2,1))
          adddose1predict(dat.now,dat.now.rma.dose1,dosescale=10)
          print(dat.now.bmd1<-addbmd1(dat.now,dat.now.rma.dose1,dosescale=10))
			bmd.results.sum <- rbind(bmd.results.sum,getbmd.results.sum(dat.now.bmd1,b.lab="Highest Doses-Linear in dose10"))
          mtext("Linear model",side=3,line=-1,cex=0.8)
          dodoseplot(dat.now,effectlab=effectlab,logxlim=c(-2,1))
          adddose2predict(dat.now,dat.now.rma.dose2,dosescale=10)
          print(dat.now.bmd2<-addbmd2(dat.now,dat.now.rma.dose2,dosescale=10))
			bmd.results.sum <- rbind(bmd.results.sum,getbmd.results.sum(dat.now.bmd2,b.lab="Highest Doses-LinearQuadratic in dose10"))
          mtext("Linear-quadratic model",side=3,line=-1,cex=0.8)
          par(mfrow=c(1,1))
          # dev.off()
```


```{r write results}

write.csv(results.sum,file="PBDE.meta-results-v2.csv")
write.csv(bmd.results.sum,file="PBDE.meta-bmd-results-v2.csv")

```


```{r Morris Latency last trial animal by chemical, fig.width=6.5, fig.height = 8, echo=FALSE, dpi=300}
##### Morris Latency_animal_forest_dose


dat<-Morris.dat.trans
for (chem.now in unique(dat$chemical)) {
results.sum <- data.frame()
bmd.results.sum<-data.frame()
  
effectlab<-"Latency last trial log(Ratio of mean)x100"
dat$slab <- paste(dat$study.name,"\n  ",dat$chemical,dat$animal.group.name)
      # All doses for each study
      dat.now <- subset(dat,chemical==chem.now & !is.na(stdev) & !is.na(response))
      studylist <- unique(dat.now$study.name)
      nstudy<-length(studylist)
      if (dim(dat.now)[1]>1) {
            # Intercept only
        print("Overall Effect---------------------")
        dat.now.rma <- doforest(dat.now,effectlab=effectlab,
                                main=paste(chem.now,"All Doses"),cex=0.7)
		results.sum <- rbind(results.sum,getresults.sum(dat.now.rma,b.lab="Overall",bindx=1))
        #        print(dat.now.rma.l1o<-leave1out(dat.now.rma))
        # Leave one out
        print("Leave one out---------------------")
        studylist <- unique(dat.now$study.name)
        nstudy<-length(studylist)
        dat.loo.rma.list<-list()
        if (nstudy>2) { 
          for (n in 1:nstudy) {
            study.now<-studylist[n]
            dat.loo <- subset(dat.now,study.name != study.now)
            print(dat.loo.rma.list[[n]]<-rma(yi,vi=vi,slab=slab,data=dat.loo))
			results.sum <- rbind(results.sum,getresults.sum(dat.loo.rma.list[[n]],
				b.lab=paste("Overall minus",study.now),bindx=1))
            #print(predict(dat.loo.rma.list[[n]]))
            # forest(dat.loo$yi.back,ci.lb=dat.loo$ci.lb.back,ci.ub=dat.loo$ci.ub.back,slab=dat.loo$slab,
            #                             xlab=effectlab,xlim=c(xmin,xmax),at=xtx,
            #                             ilab=cbind(dat.loo$dose),
            #                             ilab.xpos=dosepos,cex=cex.lab,
            #                             rows=rnums,ylim=c(-1,ymax),
            #                             digits=2
            # )
          }
        }
        # Intercept and log10(dose) - test for trend
        print("Linear in log10(dose)---------------------")
        dat.now.rma.dose <- doforestdose(dat.now,ylim=c(-1.5,dim(dat.now)[1]+3),
                                         effectlab=effectlab,
                                         main=paste(chem.now,"All Doses"),cex=0.7)
		results.sum <- rbind(results.sum,getresults.sum(dat.now.rma.dose,b.lab="Trend in log10(dose)",bindx=2))
        # Linear - no intercept - for estimating BMD
        print("Linear in dose---------------------")
        dat.now.rma.dose1 <- doforestdose1(dat.now,ylim=c(-2.5,dim(dat.now)[1]+3),
                                           effectlab=effectlab,
                                           main=paste(chem.now,"All Doses"),cex=0.7)
		results.sum <- rbind(results.sum,getresults.sum(dat.now.rma.dose1,b.lab="Linear in dose10",bindx=1))
        # Quadratic - no intercept - for estimating BMD
        print("Linear or LinearQuadratic in dose---------------------")
        dat.now.rma.dose2 <- doforestdose2(dat.now,ylim=c(-2.5,dim(dat.now)[1]+3),
                                           effectlab=effectlab,
                                           main=paste(chem.now,"All Doses"),cex=0.7)
		results.sum <- rbind(results.sum,getresults.sum(dat.now.rma.dose2,b.lab="LinearQuadratic in dose10",bindx=1))
		results.sum <- rbind(results.sum,getresults.sum(dat.now.rma.dose2,b.lab="LinearQuadratic in dose10",bindx=2))
        par(mfrow=c(3,1))
        dodoseplot(dat.now,effectlab=effectlab,logxlim=c(-2,1))
        adddosepredict(dat.now,dat.now.rma.dose)
        mtext("Log-linear model",side=3,line=-1,cex=0.8)
        dodoseplot(dat.now,effectlab=effectlab,logxlim=c(-2,1))
        adddose1predict(dat.now,dat.now.rma.dose1,dosescale=10)
        print(dat.now.bmd1<-addbmd1(dat.now,dat.now.rma.dose1,dosescale=10))
		bmd.results.sum <- rbind(bmd.results.sum,getbmd.results.sum(dat.now.bmd1,b.lab="Linear in dose10"))
        mtext("Linear model",side=3,line=-1,cex=0.8)
        dodoseplot(dat.now,effectlab=effectlab,logxlim=c(-2,1))
        adddose2predict(dat.now,dat.now.rma.dose2,dosescale=10)
        print(dat.now.bmd2<-addbmd2(dat.now,dat.now.rma.dose2,dosescale=10))
		bmd.results.sum <- rbind(bmd.results.sum,getbmd.results.sum(dat.now.bmd2,b.lab="LinearQuadratic in dose10"))
        mtext("Linear-quadratic model",side=3,line=-1,cex=0.8)
        par(mfrow=c(1,1))
        # dev.off()

        
      # highest dose for each study
      print(paste(effectlab,"Highest Dose"))
      dat.now <- subset(dat.now,high_dose==dose)
      if (dim(dat.now)[1]>1) {
        # Intercept only
        # pdf(file=paste(phthalate.now,"HighestDose",species.now,"AGD.V5.pdf",sep="."),height=8,width=6.5)
        print("Highest Dose Overall Effect---------------------")
        dat.now.rma <- doforest(dat.now,effectlab=effectlab,
                                main=paste(chem.now,"Highest Doses"),cex=0.7)
		results.sum <- rbind(results.sum,getresults.sum(dat.now.rma,b.lab="Highest Doses-Overall",bindx=1))
		
		 # Leave one out
        print("Highest Dose Leave one out---------------------")
        studylist <- unique(dat.now$study.name)
        nstudy<-length(studylist)
        dat.loo.rma.list<-list()
        if (nstudy>2) { 
          for (n in 1:nstudy) {
            study.now<-studylist[n]
            dat.loo <- subset(dat.now,study.name != study.now)
            print(dat.loo.rma.list[[n]]<-rma(yi,vi=vi,slab=slab,data=dat.loo))
			results.sum <- rbind(results.sum,getresults.sum(dat.loo.rma.list[[n]],
				b.lab=paste("Overall minus",study.now),bindx=1))
            #print(predict(dat.loo.rma.list[[n]]))
            # forest(dat.loo$yi.back,ci.lb=dat.loo$ci.lb.back,ci.ub=dat.loo$ci.ub.back,slab=dat.loo$slab,
            #                             xlab=effectlab,xlim=c(xmin,xmax),at=xtx,
            #                             ilab=cbind(dat.loo$dose),
            #                             ilab.xpos=dosepos,cex=cex.lab,
            #                             rows=rnums,ylim=c(-1,ymax),
            #                             digits=2
            # )
          }
        }
        #        print(dat.now.rma.l1o<-leave1out(dat.now.rma))
        # Linear - no intercept - for estimating BMD
        print("Highest Dose Linear or LinearQuadratic in dose---------------------")
        dat.now.rma.dose1 <- doforestdose1(dat.now,ylim=c(-2.5,dim(dat.now)[1]+3),
                                           effectlab=effectlab,
                                           main=paste(chem.now,"Highest Doses"),cex=0.7)
		results.sum <- rbind(results.sum,getresults.sum(dat.now.rma.dose1,b.lab="Highest Doses-Linear in dose10",bindx=1))
			bmd.results.sum <- rbind(bmd.results.sum,getbmd.results.sum(dat.now.bmd1,b.lab="Highest Doses-Linear in dose10"))
          mtext("Linear model",side=3,line=-1,cex=0.8)
          dodoseplot(dat.now,effectlab=effectlab,logxlim=c(-2,1))
          adddose2predict(dat.now,dat.now.rma.dose2,dosescale=10)
          print(dat.now.bmd2<-addbmd2(dat.now,dat.now.rma.dose2,dosescale=10))
		    if (dim(dat.now)[1]>2) {
        # Intercept and log10(dose) - test for trend
          print("Highest Dose Linear in log10(dose)---------------------")
          dat.now.rma.dose <- doforestdose(dat.now,ylim=c(-1.5,dim(dat.now)[1]+3),
                                           effectlab=effectlab,
                                           main=paste(chem.now,"Highest Doses"),cex=0.7)
		results.sum <- rbind(results.sum,getresults.sum(dat.now.rma.dose,b.lab="Highest Doses-Trend in log10(dose)",bindx=2))
          # Quadratic - no intercept - for estimating BMD
          print("Highest Dose LinearQuadratic in dose---------------------")
          dat.now.rma.dose2 <- doforestdose2(dat.now,ylim=c(-2.5,dim(dat.now)[1]+3),
                                             effectlab=effectlab,
                                             main=paste(chem.now,"Highest Doses"),cex=0.7)
		results.sum <- rbind(results.sum,getresults.sum(dat.now.rma.dose2,b.lab="Highest Doses-LinearQuadratic in dose10",bindx=1))
		results.sum <- rbind(results.sum,getresults.sum(dat.now.rma.dose2,b.lab="Highest Doses-LinearQuadratic in dose10",bindx=2))
          par(mfrow=c(3,1))
          dodoseplot(dat.now,effectlab=effectlab,logxlim=c(-2,1))
          adddosepredict(dat.now,dat.now.rma.dose)
          mtext("Log-linear model",side=3,line=-1,cex=0.8)
          dodoseplot(dat.now,effectlab=effectlab,logxlim=c(-2,1))
          adddose1predict(dat.now,dat.now.rma.dose1,dosescale=10)
          print(dat.now.bmd1<-addbmd1(dat.now,dat.now.rma.dose1,dosescale=10))
			bmd.results.sum <- rbind(bmd.results.sum,getbmd.results.sum(dat.now.bmd2,b.lab="Highest Doses-LinearQuadratic in dose10"))
          mtext("Linear-quadratic model",side=3,line=-1,cex=0.8)
          par(mfrow=c(1,1))
          # dev.off()
		    }
      }
write.csv(results.sum,file=paste(chem.now,"PBDE.meta-results-v2.csv",sep="."))
write.csv(bmd.results.sum,file=paste(chem.now,"PBDE.meta-bmd-results-v2.csv",sep="."))
      }
}
```
